A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2020-02-17 paper/main.tex
finalise regression and conclustion

The aim of this section is to explore the features that contribute to a
strategy's successful performance. The features explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_features}.

@@

APL makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage feature is calculated as the
memory size of strategy (which is specified in the strategies implementation
in the APL) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage feature is not used for probabilistic ending
tournaments. The SSE is a feature introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the features considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

@@

A second approach for evaluating the features importance in a strategy's success
is a combination of clustering and a random forest algorithm. Initially,
the strategies performances are clustered four different time based on the
following approaches:

@@

The importance that the features of Table~\ref{table:manual_features} had on
each random forest model while the performances were clustered based on the
different approaches have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highlighted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.

@@

A multivariate linear regression is also fitted to model the relationship of the
features and the normalised rank. The \(p-\)values of each feature as well as
their coefficients are given by Table~\ref{table:linear_regression}. The
coefficient of determination \(R\) squared fo each model is also report in
Table~\ref{table:linear_regression}. Based on \(R\) square the model explain 90\% of
the variability of the response data around its mean.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
\resizebox{.8\textwidth}{!}{
    \input{regression_table.tex}}
    \end{center}
    \caption{Multivariate linear regression with \(r\) as the depended variable results.}
    \label{table:linear_regression}
\end{table}

The results of the linear regression verify the results of the previous approaches.

The features $C_{r} / C_{median}$ and $C_r / C_{mean}$ have been repeatably
highlighted as importance feature in explaining the variability of the performances.
Thus the effect of these tow measures is further explored.In
Figure~\ref{fig:mean_median_std} the distributions of \(C_r / C_{\text{mean}}\)
and \(C_r / C_{\text{median}}\) are given for the winners in standard tournaments. A value of \(C_r /
C_{\text{mean}} = 1\) imply that the cooperating ratio of the winner was the
same as the mean/median cooperating ratio of the tournament. In standard tournaments, the mean
for both ratios is 1. Therefore, an effective strategy in standard tournaments
was the mean/median cooperator of its respective tournament. In comparison,
Figure~\ref{fig:mean_median_noisy} shows the distributions of the features for
the winners in noisy tournaments where the mean is at 0.67. Thereupon the winners
cooperated 67\% of the times the mean/median cooperator did. This analysis is
applied to the rest of the tournaments and the distributions are given by
Figures \ref{fig:mean_median_probend}, \ref{fig:mean_median_probend_noisy} and
\ref{fig:mean_median_overall}. In a tournament with noisy and a probabilistic
ending the winners cooperated 60\%, whereas in settings that the type of the
tournament can vary between the types considered in this work the winners
cooperated 67\% of the times the mean or median cooperator did. Finally, in
probabilistic ending tournament it has already been mentioned that defecting
strategies prevail and this result is once again confirmed in this section.

@@

In this section the effect of several features, regarding a strategy's behaviour
and the tournament in which it participated on its performance were presented.
This was done using three approaches. Correlation coefficients, a random forest
analysis and a linear regression. The results of these are summarised in the following section.

@@

This manuscript has explored the performance of \numberofstrategies strategies of the Iterated
Prisoner's Dilemma in a large number of computer tournaments. The results of
the analysis demonstrated that, although for specific tournament types such as
standard and probabilistic ending tournaments, dominant types of strategies exist
there is not a single dominant strategy in an IPD competition if the environments
vary. Moreover, a strategy with a theory of mind should aim to adapt its behaviour
based on the mean and median cooperators and should in general not be too
cooperative.

The \numberofstrategies strategies used in this manuscript have been mainly for
the literature, and they have been accessible due to an open source software
called the Axelrod-Python library. The software was used to generate a total of
\numberofalltournaments computer tournaments results with different number of
strategies and different participants each time. The data collection was
described in Section~\ref{section:data_collection}. In
Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments there are not strategies that can guarantee winning.
However, if the probability of noisy was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.
Overall, the top ranked strategies differed from one tournament type to
another and the mechanism behind the winning strategies were all different.
Even strategies designed to perform well in one setting were demonstrated to be
better in others.

>>>>

The aim of this section is to explore the features that contribute to a
strategy's successful performance. The features explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. A summary of the features is given in
Table~\ref{table:manual_features}.

@@

APL makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage feature is calculated as the
memory size of strategy (which is specified in the strategies implementation
in the APL) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. If the memory size of a strategy is infinite then memory usage is 1.
For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage feature is not used for probabilistic ending
tournaments. The SSE is a feature introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the features considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

The performances are evaluated based on different methods.

@@

The second method to evaluate the features importance in a strategy's success
is a combination of a clustering task and a random forest algorithm.
Initially the performances are clustered into different clusters
based on them being successful or not. The performances are clustered into
successful and unsuccessful clusters based on 4 different approaches.
More specifically:

@@

The importance that the features of Table~\ref{table:manual_features} had on
each random forest model are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highlighted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.

@@

Finally, a multivariate linear regression is fitted to model the relationship of
Table's~\ref{table:manual_features} and the normalised rank. The \(p-\)values of
each feature as well as their coefficients are given by
Table~\ref{table:linear_regression}. The coefficient of determination, \(R\)
squared, for each model is also reported in Table~\ref{table:linear_regression}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
\resizebox{.8\textwidth}{!}{
    \input{regression_table.tex}}
    \end{center}
    \caption{Results of multivariate linear regressions with \(r\) as the depended variable.}
    \label{table:linear_regression}
\end{table}

The results of the linear regression verify the results of the previous
approaches. The only features that do no have a significant effect on the
models are the number of repetitions and memory usage in noisy tournaments,
and the number of turns, number of repetitions and thr probability \(p_e\) in
probabilistic ending tournaments. Over all the models, the features that are the
most important predictors are $C_r$, $C_r$ / $C_{max}$ and $C_r$ / $C_{mean}$.

The features $C_{r} / C_{median}$ and $C_r / C_{mean}$ have been repeatably
highlighted as importance features in explaining the variability of the performances.
Thus, the effect of these two measures is further explored. In
Figure~\ref{fig:mean_median_std} the distributions of \(C_r / C_{\text{mean}}\)
and \(C_r / C_{\text{median}}\) are given for the winners in standard tournaments. A value of \(C_r /
C_{\text{mean}} = 1\) imply that the cooperating ratio of the winner was the
same as the mean/median cooperating ratio of the tournament. In standard tournaments, the mean
for both ratios is 1. Therefore, an effective strategy in standard tournaments
was the mean/median cooperator of its respective tournament. In comparison,
Figure~\ref{fig:mean_median_noisy} shows the distributions of the features for
the winners in noisy tournaments where the mean is at 0.67. Thereupon the winners
cooperated 67\% of the times the mean/median cooperator did. This analysis is
applied to the rest of the tournaments and the distributions are given by
Figures \ref{fig:mean_median_probend}, \ref{fig:mean_median_probend_noisy} and
\ref{fig:mean_median_overall}. In a tournament with noisy and a probabilistic
ending the winners cooperated 60\%, whereas in settings that the type of the
tournament can vary between the types considered in this work the winners
cooperated 67\% of the times the mean or median cooperator did. Finally, in
probabilistic ending tournament it has already been mentioned that defecting
strategies prevail and this result is once again confirmed in this section.

@@

In this section the effect of several features, regarding a strategy's behaviour
and the tournament in which it participated on its performance were presented.
This was done using three methods. Correlation coefficients, a random forest
analysis and a linear regression. The results of these are summarised in the following section.

@@

This manuscript has explored the performance of \numberofstrategies strategies of the Iterated
Prisoner's Dilemma in a large number of computer tournaments. The results of
the analysis demonstrated that, although for specific tournament types such as
standard, probabilistic ending tournaments and noisy tournaments with \(p_n <0.1\), dominant types of strategies exist
there is not a single dominant strategy in an IPD competition if the environments
vary. Moreover, a strategy with a theory of mind should aim to adapt its behaviour
based on the mean and median cooperators and should in general not be too
cooperative.

The \numberofstrategies strategies used in this manuscript have been mainly for
the literature, and they have been accessible due to an open source software
called the Axelrod-Python library. The software was used to generate a total of
\numberofalltournaments computer tournaments results with different number of
strategies and different participants each time. The data collection was
described in Section~\ref{section:data_collection}. In
Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments there are not strategies that can guarantee winning.
However, if the probability of noisy was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.
Overall, the top ranked strategies differed from one tournament type to
another and the mechanism behind the winning strategies were all different, with some
exceptions which include trained (Evolved) and Meta strategies.
